CALVIS: chest, waist and pelvis circumference from 3D human body meshes as ground truth for deep learning

### 2019-10-27 PaperRMFI2019/samplepaper.tex
add pictures

\begin{document}
%
\title{CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez-Tejeda and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate chest, waist and 
pelvis circumference from 3D human body meshes. Our motivation is to use 
this data as ground truth for learning with convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
$\textit{manually}$ these anthropometrical measurements directly from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast our method can be used on 
3D meshes automatically. We conduct two experiments. 
In the first experiment we synthesize 10 human body meshes. Then we apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively. We observe that the measurements can indeed be used 
to estimate the shape of a person. The second experiment asses the 
plausibility of our approach where we use the calculated human dimensions as 
a ground truth to train a small CNN. After 
having trained the network with 
our data, we demonstrate that learning converges, achieving $x$ percent 
prediction error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

The problem of estimating chest, waist and 
pelvis circumference having only an image of a person is a challenging problem 
and involve recognizing the corresponding location of these dimension in the 
body and estimating their circumference. Additionally, the setting is an  
under-constrained  (or inverse) problem. Information 
gets lost when a camera is 
used to capture the human body in 3D space to 'render' a 2D image.

@@

%------------------------------------------------------------------------
\section{Approach}\label{sec:approach}
While human body 
meshes are traditionally acquired by specialized 3D scanners, we use a 
generative model to synthesize them, which we briefly review in 
subsection \ref{subsec:smpl_model}. To calculate the actual human body 
dimensions, we first need to formalize the 
notion of chest, waist and pelvis of a 3D human body mesh. Our strategy consist 
of segmentating the mesh in five regions, 
three of them of interest, which we 
discuss in subsection \ref{subsec:three_regions}. Finally, within these regions 
we identify the intended dimensions employing a human body mesh signature 
(HBMS), 
defined in subsection \ref{subsec:hbm_signature}.

@@

The SMPL model is at its top level a \textbf{skinned articulated 
model}, i.e., 
consists of a 
surface mesh $\mathcal{M}$ that mimics the skin and a corresponding skeleton 
$\mathbf{J}$. The mesh $\mathcal{M} \in \mathbb{R}^{3N}$, which is a boundary 
representation stores 
both the body geometry (vertex position) and topology (how the vertices are 
connected). The skeleton $\mathbf{J}$ is defined by its joints location in 3D 
space $j_i \in \mathbb{R}^3$  and its kinematic tree. Two 
connected 
joints define a `bone'. Moreover, a child bone is rotated relative to 
its 
connected parent. The pose $\vec{\theta}$ is the specification of every bone 
rotation plus an orientation for the root bone.

@@

As mentioned above $\mathbf{\bar{T}} \in \mathbb{R}^{3N}$  is the mean template 
shape. The set of blend weights $\mathcal{W} \in \mathbb{R}^{N \times K}$ 
represents 
how much the rotation of 
skeleton bones affects
the vertices. In addition, the matrices $\mathcal{S}$ and $\mathcal{P}$ define 
respectively linear functions $B_s(\vec{\beta}; \mathcal{S}): 
\mathbb{R}^{|\vec{\beta}|} \to 
\mathbb{R}^{3N}$ and $B_p(\vec{\theta}; \mathcal{P}): 
\mathbb{R}^{|\vec{\theta}|} \to 
\mathbb{R}^{3N}$ that are used to deform $\mathbf{\bar{T}}$; and the function 
$\mathcal{J}: \mathbb{R}^{|\vec{\beta}|} \to \mathbb{R}^{3K}$ predicts skeleton 
rest joint locations from vertices in the rest 
pose. A new mesh $\mathcal{M}_{new}$ can be then generated

@@

Let us consider a human body mesh $\mathcal{M}$. Our method requires that 
$\mathcal{M}$ is standing with arms raised 
parallel to the 
ground at shoulder height at a $90^\circ$ angle (SMPL normalized pose
$\vec{\theta}^*$). Additionally, we assume that the mesh 
has LSA orientation, e.g., x, y and z axis are positively directed from 
right-to-left, inferior-to-superior and posterior-to-anterior, respectively. If 
the mesh has another orientation we can always apply transformations to 
LSA-align it.

@@

We observe that the chest circumference is usually measure bellow the armpits, 
also known as axilla, but above the natural waist. Similarly, the pelvis 
circumference is measured often around the hips. This observation suggests that 
we should consider segmenting the mesh in meaningful regions. Moreover, we can 
use the skeleton when determining the region boundaries. For 
example, there is consensus regarding the axilla 
definition as the area of the human body underneath the shoulder joint 
(\cite{MeSH.axilla}, \cite{FMA.axilla}, \cite{TA.axilla}). Therefore, we can 
use the shoulder joint as a hint to locate the axilla and establish a upper 
bound for the chest region. More generally, we can use the 
skeleton joints defined in $\mathbf{J}$ to identify the regions boundaries.

Let the axilla location be the point $p_a = (p^x_a, p^y_a, p^z_a)$ and 
$j_{nw}$, $j_{pl}$, 
$j_{lh} \in \mathbf{J}$ be the natural waist, pelvis and left hip joints, 
respectively.

@@

\begin{align}\label{eq:pr}
\mathcal{PR} = \big(v^y_i \, | \, j^y_{pl} > v^y_i \geq j^y_{lh}\big)
\end{align}

@@

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla. What we want is a point $p_a$ in the mesh at the proper location 
under 
the arms. One point suffices because we require only a reference $p^y_a$ along 
the 
y-axis to define the chest region according to equation \ref{eq:cr}. Therefore, 
we focus on the right shoulder joint $j_{rs} \in \mathbf{J}$. Since the axilla 
is related to 
the shoulder joint, we can cast a ray $\boldsymbol{\iota}$ from it in direction 
to the middle left 
inferior edge of the bounding box outside the mesh at point

\begin{align}
p_c = \bigg(v^x_{min}, v^y_{min}, v^z_{min} + 
\frac{v^z_{max} - v^z_{min}}{2}\bigg).
\end{align}

Here $v^x_{min}, v^y_{min}, v^z_{min}$ are vertex minimum $x$,$y$ and $z$ 
coordinates, respectively; analog is $v^z_{max}$ the maximum vertex $z$ 
coordinate. The ray 
$\iota$ is then determined by points $j_{rs}$ (inside the mesh) and $p_b$ 
(outside the mesh). Therefore, the ray $\iota$ intersects the mesh at one point 
$p_c$ (Figure \ref{fig:axilla-recognition}), which we consider the axilla 
center. Furthermore, we can consider an strategy to increase the method 
robustness against the large variability of the human body. Since our method 
relies on one point to establish the chest upper bound, we identify the 10 
nearest neighbors in the set of vertices $\mathcal{V}$ to $p_c$ and define 
$p_a$ to be the 
vertex with the smallest 
y coordinate of the neighbors.

@@

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{axilla_recognition_final.png}
	\end{center}
	\caption{Axilla recognition based on the skeleton joint. We cast a ray from 
	the left shoulder joint $j_{ls}$ in direction to the middle left 
		inferior edge of the bounding box (left). The ray intersects the mesh 
		at point $p_a$ that we consider the axilla center. We plot the 10 
		nearest vertices to $p_a$ (right). As it can be observed, the axilla is 
		properly recognized.}
	\label{fig:axilla-recognition}
\end{figure}

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Intuitively, we would like to measure the chest circumference below the arms 
at the \textit{widest} part of the torso and the waist circumference at the 
\textit{narrowest} part beneath the chest but above the hips. This intuition is 
compliant 
with prior body dimensions standardized definition, for example, the European 
standard EN 13402-1 \cite{en13402-1}. Similarly, the pelvis 
circumference is measured often around the \textit{widest} part of hips and 
buttocks. The general idea is to cast the chest, waist and pelvis circumference 
estimation problem as an optimization problem. If we are able to establish a 
function that measures geodesics on the mesh, we can search for its maximum, 
e.g., chest in the chest region.

We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves\cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the intersection boundary length as
\begin{equation}\label{eq:boundary_length}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, q) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

\begin{figure}[h!]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=0.5\linewidth]{subject_2_with_cutting_planes_xrays.png}
	\end{center}
	\caption{3D mesh iterative slicing. We 
		can use a plane $\boldsymbol{\pi_j}$ parallel (with normal 
		$\vec{n}_{||}$) to the 
		floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j 
		\in 
		\mathbb{R}$ along the y-axis. By varying point $q_j$ (we show here 40), 
		we can obtain 
		slices.}
	\label{fig:subjects_planes_xrays}
\end{figure}

Starting 
from the 
top $q_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters along the y-axis until we reach the 
bounding box bottom 
$q_b$. Figure \ref{fig:subjects_planes_xrays} shows 40 of these 
slides. Next, we assemble the mesh sliding vector $\vec{\mathcal{L}}$. The 
elements of 
this vector are slice points

\begin{align}\label{eq:mesh_slice_vector}
\vec{\mathcal{L}}(\mathcal{M}, \vec{n}_{||}, m) = \left[ q_t, q_t+m, \cdots, 
q_b \right]
\end{align}

Finally, we apply equation \ref{eq:boundary_length} to every element of 
$\mathcal{L}$ in equation \ref{eq:mesh_slice_vector}. In other words, for every 
slice at point 
$q_j$, we compute $\mathcal{BL}(q_j)$. Here we drop from the notation mesh 
$\mathcal{M}$ and plane normal $\vec{n}_{||}$ because they remain constant. We 
then can 
define the \textbf{human body mesh signature} $\mathcal{MS}: 
\vec{\mathcal{L}} \to \mathbb{R}$ that maps every slice at point $q_j$ to the 
corresponding boundary length $\mathcal{BL}(q_j)$.

@@

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(\mathcal{M}, \vec{n}_{||}, m) = 
[\mathcal{BL}(q_t), \mathcal{BL}(q_t+m), \cdots, \mathcal{BL}(q_b)]
\end{align}

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:hbm_signature}
\end{figure}

@@

In general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The chest circumference $cc$ is the local maximum within the chest 
	region.
	\item The waist circumference $wc$ (not to be confused with the natural 
	waist line) is the minimum within the waist region.	
	\item The pelvis circumference $pc$ is the maximum within the pelvis region.
\end{enumerate}

\begin{align}\label{eq:cc}
cc &= \mathcal{L}_i \in \mathcal{CR}\\
\mathcal{MS}(\mathcal{M}, \vec{n}_{||}, m) &= \mathcal{L}_i \in \mathcal{CR}
\end{align}

@@

\subsection{Qualitative evaluation}\label{subsec:qualitative_eval}
Here we evaluate our method qualitatively with method one and method two.
\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{subject_6_with_cutting_planes_on_joints.eps}
	\end{center}
	\caption{Qualitative evaluation for male and female meshes. We evaluate our 
	method with method one and method two.}
	\label{fig:qualitative_eval}
\end{figure}

Figure \ref{fig:qualitative_eval} shows the $\mathcal{MS}$ of two meshes (male 
and female) for 
$m=0,001$ and plane $\boldsymbol{\pi}$ parallel to the floor (with normal $(0, 
1,0)$) using the library trimesh \cite{trimesh}. Note that the function as 
defined by equation \ref{eq:mesh_signature} is bounded and not continuous. It 
resembles a rotated silhouette of the human body and exhibits several 
\textit{extrema}.

\subsection{Calvis-Net and Learning Convergence 
Evaluation}\label{subsec:learn_conv}
This experiment demonstrates the plausibility of using calvis to obtain data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for ANN.

@@

Figure \ref{fig:learning_curve} shows that after epoch number 5 the estimation 
error decreases and the learning converges.
\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Calvis-Net.png}
	\end{center}
	\caption{Calvis-Net.}
	\label{fig:calvis_net}
\end{figure}

>>>>

\begin{document}
%
\title{CALVIS: chest, waist and pelvis circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez-Tejeda and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate \textbf{C}hest, 
w\textbf{A}ist and pe\textbf{LVIS} circumference from 3D human body meshes. Our 
motivation is to use 
this data as ground truth for training convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
these anthropometrical measurements $\textit{manually}$ from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast, our method can be used on 
3D meshes automatically. We synthesize eight human body meshes and 
apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively and observe that the measurements can indeed be used 
to estimate the shape of a person. We then asses the 
plausibility of our approach by exploiting the calculated human dimensions as 
ground truth to train a small CNN. After
having trained the network with 
our data, we achieve competitive prediction error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

The problem of estimating chest, waist and 
pelvis circumference having only an image of a person is a challenging problem 
and involve recognizing the corresponding location of these dimensions in the 
body and estimating their circumference. Additionally, the setting is an  
under-constrained  (or inverse) problem. Information 
gets lost when a camera is 
used to capture the human body in 3D space to 'render' a 2D image.

@@

%------------------------------------------------------------------------
\section{Chest, Waist and Pelvis Measurements}\label{sec:approach}
While human body 
meshes are traditionally acquired by specialized 3D scanners, we use a 
generative model to synthesize them, which we briefly review in 
subsection \ref{subsec:smpl_model}. To calculate the actual human body 
dimensions, we first need to formalize the 
notion of chest, waist and pelvis of a 3D human body mesh. Our strategy consist 
of segmentating the mesh in five regions, 
three of them of interest, which we 
discuss in subsection \ref{subsec:three_regions}. Finally, within these regions 
we identify the intended dimensions employing a human body mesh signature 
(HBMS), 
defined in subsection \ref{subsec:hbm_signature}.

@@

The SMPL model is at its top level a \textbf{skinned articulated 
model}, i.e., 
consists of a 
surface mesh $\mathcal{M}$ that mimics the skin and a corresponding skeleton 
$\mathbf{J}$. The mesh $\mathcal{M}$, which is a boundary 
representation stores 
both the body geometry (vertex position) and topology (how the vertices are 
connected). The skeleton $\mathbf{J}$ is defined by its joints location in 3D 
space $j_i \in \mathbb{R}^3$  and its kinematic tree. Two 
connected 
joints define a `bone'. Moreover, a child bone is rotated relative to 
its 
connected parent. The pose $\vec{\theta}$ is the specification of every bone 
rotation plus an orientation for the root bone.

@@

As mentioned above $\mathbf{\bar{T}} \in \mathbb{R}^{3N}$  is the mean template 
shape. The set of blend weights $\mathcal{W} \in \mathbb{R}^{N \times K}$ 
represents 
how much the rotation of 
skeleton bones affects
the vertices. In addition, the matrices $\mathcal{S}$ and $\mathcal{P}$ define 
respectively linear functions $B_s(\vec{\beta}; \mathcal{S}): 
\mathbb{R}^{|\vec{\beta}|} \to 
\mathbb{R}^{3N}$ and $B_p(\vec{\theta}; \mathcal{P}): 
\mathbb{R}^{|\vec{\theta}|} \to 
\mathbb{R}^{3N}$ that are used to deform $\mathbf{\bar{T}}$; and the function 
$\mathcal{J}: \mathbb{R}^{|\vec{\beta}|} \to \mathbb{R}^{3K}$ predicts skeleton 
rest joint locations from vertices in the rest 
pose. A new mesh $\mathcal{M}_{new}$ can be then generated using the SMPL model 
$M$

@@

Let us consider a human body mesh $\mathcal{M}$. Our method requires that 
$\mathcal{M}$ is standing with arms raised 
parallel to the 
ground at shoulder height at a $90^\circ$ angle (SMPL zero pose
$\vec{\theta}^*$). Additionally, we assume that the mesh 
has LSA orientation, e.g., x, y and z axis are positively directed from 
right-to-left, inferior-to-superior and posterior-to-anterior, respectively. If 
the mesh has another orientation we can always apply rigid transformations to 
LSA-align it.

@@

We observe that the chest circumference is usually measure below the armpits, 
also known as axilla, but above the natural waist. Similarly, the pelvis 
circumference is measured often around the hips. This observation suggests that 
we should consider segmenting the mesh in meaningful regions. Moreover, we can 
use the skeleton when determining the region boundaries. For 
example, there is consensus regarding the axilla 
definition as the area of the human body underneath the shoulder joint 
(\cite{MeSH.axilla}, \cite{FMA.axilla}, \cite{TA.axilla}). Therefore, we can 
use the shoulder joint as a hint to locate the axilla and establish a upper 
bound for the chest region. More generally, we can use the 
skeleton joints defined in $\mathbf{J}$ to identify the regions boundaries.

Let the axilla location be the point $p_a = (p^x_a, p^y_a, p^z_a)$ and 
$j_{nw}$, $j_{pl}$, 
$j_{rh} \in \mathbf{J}$ be the natural waist, pelvis and right hip joints 
(joints with identifier $\mathtt{Spine1}$, $\mathtt{Pelvis}$, $\mathtt{R\_Hip}$ 
in 
$\mathbf{J}$), respectively.

@@

\begin{align}\label{eq:pr}
\mathcal{PR} = \big(v^y_i \, | \, j^y_{pl} > v^y_i \geq j^y_{rh}\big)
\end{align}

@@

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question how we can recognize 
the axilla. What we want is a point $p_a$ in the mesh at the proper location 
under 
the arms. One point suffices because we require only a reference $p^y_a$ along 
the 
y-axis to define the chest region according to equation \ref{eq:cr}. Therefore, 
we focus on the right shoulder joint $j_{rs} \in \mathbf{J}$. Since the axilla 
is related to 
the shoulder joint, we can cast a ray $\boldsymbol{\iota}$ from it in direction 
to the middle left 
inferior edge of the bounding box outside the mesh at point

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{axilla_recognition_final.png}
	\end{center}
	\caption{Axilla recognition based on the skeleton joint. We cast a ray from 
		the left shoulder joint $j_{ls}$ in direction to the middle left 
		inferior edge of the bounding box (left). The ray intersects the mesh 
		at point $p_c$ (red) that we consider the axilla center. We plot the 10 
		nearest vertices to $p_c$ (right). Point $p_a$ is the nearest vertex 
		with smallest y coordinate. As it can be observed, the axilla is 
		properly recognized.}
	\label{fig:axilla-recognition}
\end{figure}

\begin{align}
p_c = \bigg(v^x_{min}, v^y_{min}, v^z_{min} + 
\frac{|v^z_{max} - v^z_{min}|}{2}\bigg).
\end{align}

Here $v^x_{min}, v^y_{min}, v^z_{min}$ are vertex minimum $x$,$y$ and $z$ 
coordinates, respectively; analog is $v^z_{max}$ the maximum vertex $z$ 
coordinate. The ray 
$\iota$ is then determined by points $j_{rs}$ (inside the mesh) and $p_b$ 
(outside the mesh). Therefore, the ray $\iota$ intersects the mesh at one point 
$p_c$ (Figure \ref{fig:axilla-recognition}), which we consider the axilla 
center. Furthermore, we can consider an strategy to increase the method 
robustness against the large variability of the human body. Since this approach 
relies on one point to establish the chest upper bound, we identify the 10 
nearest neighbors in the set of vertices $\mathcal{V}$ to $p_c$ and define 
$p_a$ to be the vertex with the smallest y coordinate of the neighbors.

@@

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Intuitively, we would like to measure the chest circumference below the arms 
at the \textit{widest} part of the torso and the waist circumference at the 
\textit{narrowest} part beneath the chest but above the hips. This intuition is 
compliant 
with prior body dimensions standardized definition, for example, the European 
standard EN 13402-1 \cite{en13402-1}. Similarly, the pelvis 
circumference is measured often around the \textit{widest} part of hips and 
buttocks. The general idea is to cast the chest, waist and pelvis circumference 
estimation problem as a constrained optimization problem. If we are able to 
establish a 
function that measures meaningful geodesics on the mesh, we can search for its 
maximum, 
e.g., chest in the chest region.

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=0.5\linewidth]{subject_2_with_cutting_planes_xrays.png}
	\end{center}
	\caption{3D mesh iterative slicing. We 
		can use a plane $\boldsymbol{\pi_j}$ parallel (with normal 
		$\vec{n}_{||}$) to the 
		floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j 
		\in 
		\mathbb{R}$ along the y-axis. By varying point $q_j$ (we show here 40), 
		we can obtain 
		slices.}
	\label{fig:subjects_planes_xrays}
\end{figure}

We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves\cite{yansel} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a polygonal curve consisting in segments with length $s_i$. 
Therefore, we 
can 
determine the intersection boundary length as
\begin{equation}\label{eq:boundary_length}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, q) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

Starting 
from the 
top $q_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters along the y-axis until we reach the 
bounding box bottom 
$q_b$. Figure \ref{fig:subjects_planes_xrays} shows 40 of these 
slices. Next, we assemble the mesh slicing vector $\vec{\mathcal{L}}$. The 
elements of 
this vector are slice points

\begin{align}\label{eq:mesh_slice_vector}
\vec{\mathcal{L}}(\mathcal{M}, \vec{n}_{||}, m) = \left[ q_t, q_t-m, q_t-2 
m,\cdots, 
q_b \right]
\end{align}

Finally, we apply equation \ref{eq:boundary_length} to every slice defined in
$\mathcal{L}$ in equation \ref{eq:mesh_slice_vector}. In other words, for every 
slice at point 
$q_j$, we compute intersection boundary length $\mathcal{BL}(q_j)$. Here we 
drop from the notation mesh 
$\mathcal{M}$ and plane normal $\vec{n}_{||}$ because they remain constant. We 
then can 
define the \textbf{human body mesh signature} $\mathcal{MS}: 
\vec{\mathcal{L}} \to \mathbb{R}$ that maps every slice at point $q_j$ to the 
corresponding boundary length $\mathcal{BL}(q_j)$.

@@

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(\mathcal{M}, \vec{n}_{||}, m) = 
[\mathcal{BL}(q_t), \mathcal{BL}(q_t-m), \cdots, \mathcal{BL}(q_b)]
\end{align}

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:hbm_signature}
\end{figure}

@@

In general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The chest circumference $cc$ is the local maximum within the chest 
	region.
	\item The waist circumference $wc$ (not to be confused with the natural 
	waist line based on the joint, see subsection \ref{subsec:three_regions}) 
	is the minimum within the waist region.	
	\item The pelvis circumference $pc$ is the maximum within the pelvis region.
\end{enumerate}

\begin{align}\label{eq:cc}
cc = \underset{\mathcal{BL}(q_j)}{\text{arg max}}\ \mathcal{MS} \quad 
\text{subject 
to} \quad p^y_a > q_j \geq j^y_{nw}
\end{align}

\begin{align}\label{eq:wc}
wc = \underset{\mathcal{BL}(q_j)}{\text{arg min}}\ \mathcal{MS} \quad 
\text{subject 
	to} \quad j^y_{nw} > v^y_i \geq j^y_{pl}
\end{align}

\begin{align}\label{eq:pc}
pc = \underset{\mathcal{BL}(q_j)}{\text{arg max}}\ \mathcal{MS} \quad 
\text{subject 
	to} \quad j^y_{pl} > v^y_i \geq j^y_{rh}
\end{align}

@@

\subsection{Qualitative evaluation}\label{subsec:qualitative_eval}
Here we evaluate our method qualitatively.
\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{experiment_1_results.png}
	\end{center}
	\caption{Qualitative evaluation for male and female meshes.}
	\label{fig:qualitative_eval}
\end{figure}

Figure \ref{fig:qualitative_eval} shows the calculated dimensions on eight 
subjects, male subjects no. 1-4 and female subjects no. 5-8. The black paths 
represent chest, waist and pelvis circumference (from 
top to bottom) as defined by equations \ref{eq:cc}-\ref{eq:pc}, respectively. 
We slightly rotate the meshes and add transparency for visualization purposes. 
Note that our method is able to estimate automatically the human body 
dimensions for all 
subjects. One particularity can be observed regarding the chest circumference: 
while for all male and female subjects except for subject no. 7 the chest 
circumference is 
estimated above the breast, for subject no.7 the chest circumference is 
estimated around the breast.

\subsection{Learning with Calvis-Net}\label{subsec:learning}
This experiment demonstrates the plausibility of employing calvis to obtain 
data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for CNN.

\subsubsection{Experimental setup} We annotate the CMU dataset with calvis. The 
automatic annotation process took approximately $1\,h\, 47\, min$ in an 
enhanced 
modern personal computer.

\subsubsection{Calvis-CMU Dataset} We synthesize x images of every 
person using the package Blender. To increase the realism of these images we 
adopt a yansel lighting model. This makes sense because it resembles the way 
how a person is standing in a room when she/he goes to the tailor for her/his 
body dimensions to be taken.

\subsubsection{Synthetic images as input} We synthesize x images of every 
person using the package Blender. To increase the realism of these images we 
adopt a yansel lighting model. This makes sense because it resembles the way 
how a person is standing in a room when she/he goes to the tailor for her/his 
body dimensions to be taken.

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Calvis-Net.png}
	\end{center}
	\caption{Calvis-Net.}
	\label{fig:calvis_net}
\end{figure}

@@

Figure \ref{fig:learning_curve} shows that after epoch number 5 the estimation 
error decreases and the learning converges.

We achieve mean absolute validation error of 0.061698 m plus +-.
